Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-07-25 exp1/iswc2013.tex
table and figure placement
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@626 43ef1190-90fd-4dd7-a769-d0121f02b80e
table and figure placement
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@627 43ef1190-90fd-4dd7-a769-d0121f02b80e
Under/overfull boxes
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@628 43ef1190-90fd-4dd7-a769-d0121f02b80e
draft subtitle
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@629 43ef1190-90fd-4dd7-a769-d0121f02b80e
Comment out draft
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@630 43ef1190-90fd-4dd7-a769-d0121f02b8

\begin{document}

\maketitle

@@

In an experiment, there are several factors, and one may run the
experiment by measuring the response variable for every combination of
levels. In a two-level experiment, this will result in $2^n$ runs,
where $n$ is the number of factors. This is called a \emph{full
  factorial experiment}. This is done in Section~\ref{sec:full}.
%TODO {Simple 8 x 2 overview table?}

@@

Common experiments include comparing several different
implementations. However, we focus on a slightly different problem: We
would like to compare the same implementation before and after some
change has been made, typically with the intention to enhance the
performance of the system. There are two reasons for this: One is that
there are currently many SPARQL implementations available, also Free
Software ones that can be modified by anyone, and the other is to
constrain the scope of the study to comparing just two different
things.

@@

  \begin{center}\small
    \begin{tabular}{|@{~~}p{0.40\textwidth}|@{~~}p{0.53\textwidth}|}
      \hline
      \textbf{Level 1} & \textbf{Level 2} \\[-2ex] 
      \verbatiminput{experiment/BGPComp-1}
      & \verbatiminput{experiment/BGPComp-2}\\
      \hline
    \end{tabular}
  \end{center}

@@

\begin{center}\small
  \begin{tabular}{|@{~~}p{0.40\textwidth}|@{~~}p{0.53\textwidth}|}
    \hline
    \textbf{``Union''} & \textbf{``Lang''} \\[-2ex]
    \verbatiminput{experiment/Union-2} &
    \verbatiminput{experiment/Lang-2} \\
    \hline
    \textbf{``Range''} & \textbf{``Optional''} \\[-2ex]
    \verbatiminput{experiment/Range-2} & 
    \verbatiminput{experiment/Optional-2}\\
    \hline
  \end{tabular}
\end{center}
These fragments has been carefully designed for illustrative utility,
as well as suitable selectivity. They are not themselves important,
they only serve to illustrate certain examples of possible
factors. When the experiment is run, they are combined to yield a
query, which is then sent to a SPARQL endpoint. The SPARQL endpoint is
itself set up using 4store (see \cite{harris20094store}) version
1.1.5.

@@

\begin{table}[h!]
\begin{center}
\caption{$p$-values for different parts of the experiment}\label{tab:pvaluesfull}
\begin{tabular}{cccccl}
  \hline
``TripleC'' & ``BGPComp'' & ``Lang'' & ``Union'' & ``Optional'' & $p$ \\ 
  \hline
  1 & 1 & 1 & 1 & 1 & 0.012 \\ 
  1 & 1 & 1 & 1 & 2 & $1.4 \cdot 10^{-09}$ \\ 
  1 & 1 & 1 & 2 & 1 & $3.1 \cdot 10^{-09}$ \\ 
  1 & 1 & 1 & 2 & 2 & $6.1 \cdot 10^{-11}$ \\ 
  1 & 1 & 2 & 1 & 1 & $3.3 \cdot 10^{-06}$ \\ 
  1 & 1 & 2 & 1 & 2 & $2.7 \cdot 10^{-09}$ \\ 
  1 & 1 & 2 & 2 & 1 & $2 \cdot 10^{-06}$ \\ 
  1 & 1 & 2 & 2 & 2 & $3.6 \cdot 10^{-10}$ \\ 
  1 & 2 & 1 & 1 & 1 & 0.014 \\ 
  1 & 2 & 1 & 1 & 2 & $1.2 \cdot 10^{-10}$ \\ 
  1 & 2 & 1 & 2 & 1 & $2.8 \cdot 10^{-14}$ \\ 
  1 & 2 & 1 & 2 & 2 & $4.1 \cdot 10^{-15}$ \\ 
  1 & 2 & 2 & 1 & 1 & $2.1 \cdot 10^{-05}$ \\ 
  1 & 2 & 2 & 1 & 2 & $2.7 \cdot 10^{-07}$ \\ 
  1 & 2 & 2 & 2 & 1 & 0.0072 \\ 
  1 & 2 & 2 & 2 & 2 & $1.6 \cdot 10^{-05}$ \\ 
  2 & 1 & 1 & 1 & 1 & 0.28 \\ 
  2 & 1 & 1 & 1 & 2 & $3 \cdot 10^{-07}$ \\ 
  2 & 1 & 1 & 2 & 1 & $3.3 \cdot 10^{-07}$ \\ 
  2 & 1 & 1 & 2 & 2 & $1.7 \cdot 10^{-08}$ \\ 
  2 & 1 & 2 & 1 & 1 & 0.0023 \\ 
  2 & 1 & 2 & 1 & 2 & $6.5 \cdot 10^{-07}$ \\ 
  2 & 1 & 2 & 2 & 1 & 0.00032 \\ 
  2 & 1 & 2 & 2 & 2 & $1.3 \cdot 10^{-06}$ \\ 
  2 & 2 & 1 & 1 & 1 & 0.013 \\ 
  2 & 2 & 1 & 1 & 2 & $1 \cdot 10^{-11}$ \\ 
  2 & 2 & 1 & 2 & 1 & $3.8 \cdot 10^{-11}$ \\ 
  2 & 2 & 1 & 2 & 2 & $4.1 \cdot 10^{-15}$ \\ 
  2 & 2 & 2 & 1 & 1 &   1 \\ 
  2 & 2 & 2 & 1 & 2 & $2.3 \cdot 10^{-05}$ \\ 
  2 & 2 & 2 & 2 & 1 &   1 \\ 
  2 & 2 & 2 & 2 & 2 & 0.99 \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{figure}[h!]
  \centerline{%
    \includegraphics[width=.9\textwidth]{frac32normal.pdf}}
  \caption{A normal plot of the Fractional Factorial Experiment with
    32 runs. The labelled points are considered significant by using
    the Lenth criterion at a level
    $\alpha=0.05$.}
  \label{fig:frac32normal}
\end{figure}

@@

\begin{figure}[h!]
  \centerline{%
    \includegraphics[width=0.9\textwidth]{frac64normal.pdf}}
  \caption{A normal plot of the Fractional Factorial Experiment with
    64 runs. The labelled points are considered significant by using
    the Lenth criterion. Triangles correspond to a level $\alpha=0.15$
    and crosses at $\alpha=0.05$. The use of different symbols have
    been added to FrF2 by the authors.}\label{fig:frac64normal}
\end{figure}

@@

\begin{figure}[h!]
  \centerline{%
  \includegraphics[width=.9\textwidth]{frac64hugenormal.pdf}}
  \caption{A normal plot of the Fractional Factorial Experiment with
    64 runs where the level 2 of ``TripleC'' is 15 MTriples. The
    labelled points are considered significant by using the Lenth
    criterion at a level $\alpha=0.05$.}\label{fig:frac64hugenormal}
\end{figure}

@@

%\bibliographystyle{plain}
%\bibliographystyle{abbrv}
%\bibliographystyle{jbact}
\bibliographystyle{splncs03}
\bibliography{practical,selectivity,federation,benchmarks,optimization,experimentsperformance,specs}

\end{document}

>>>>

%\subtitle{---Draft submitted to ISWC 2013---}

\begin{document}
\maketitle

@@

In an experiment, there are several factors, and one may run the
experiment by measuring the response variable for every combination of
levels. In a two-level experiment, this will result in $2^n$ runs,
where $n$ is the number of factors. This is called a \emph{full
  factorial experiment}. This is done in Section~\ref{sec:full}.

@@

Common experiments include the comparison of several different
implementations. However, we focus on a slightly different problem: We
would like to compare the same implementation before and after some
change has been made, typically with the intention to enhance the
performance of the system. There are two reasons for this: One is that
there are currently many SPARQL implementations available, also Free
Software ones that can be modified by anyone, and the other is to
constrain the scope of the study to comparing just two different
things.

@@

  \begin{center}\small
    \begin{tabular}{|@{~~}p{0.40\textwidth}|@{~~}p{0.53\textwidth}|}
      \hline
      \textbf{Level 1} & \textbf{Level 2} \\[-2ex] 
      \begin{verbatim}
?s  rdfs:label ?l1 ;
    ?p1 ?o1 .
?o1 dbo:populationTotal ?l2 .
      \end{verbatim}
      & 
      \begin{verbatim}
?s  rdfs:label ?l1 ;
    ?p1 ?o1 .
?o1 dbo:populationTotal ?l2 .
?s foaf:page ?o2 ;
   dbpprop:subdivisionName ?o3 .
?o3 skos:subject ?o4 ;
   dbpprop:seat ?o5 ;
   a ?c1 .\end{verbatim} 
\\      \hline
    \end{tabular}
  \end{center}

@@

\begin{center}\small
  \begin{tabular}{|@{~~}p{0.40\textwidth}|@{~~}p{0.53\textwidth}|}
    \hline
    \textbf{``Union''} & \textbf{``Lang''} \\[-2ex]
    \begin{verbatim}
{
  ?o1 dbpprop:longd ?long ;
      dbpprop:latd  ?lat .
} UNION {
  ?o1 geo:long ?long ;
      geo:lat ?lat .
}\end{verbatim} 
&
\begin{verbatim}
FILTER langMatches( lang(?l1), "fr" ) 
\end{verbatim} 
\\    \hline
    \textbf{``Range''} & \textbf{``Optional''} \\[-2ex]
    \begin{verbatim}
FILTER (?l2 > 800000)\end{verbatim} 
 & 
    \begin{verbatim}
OPTIONAL { ?o1 foaf:homepage ?o6 . }\end{verbatim} 
\\    \hline
  \end{tabular}
\end{center}
These fragments has been carefully designed for illustrative utility,
as well as suitable selectivity. They are not themselves important,
they only serve to illustrate certain examples of possible
factors. When the experiment is run, they are combined to yield a
query, which is then sent to a SPARQL endpoint. The SPARQL endpoint is
itself set up using 4store (see \cite{harris20094store}) version
1.1.5.

@@

\begin{table}[ht!]
\begin{center}
\caption{$p$-values for different parts of the experiment}\label{tab:pvaluesfull}
\begin{tabular}{cccccl}
  \hline
``TripleC'' & ``BGPComp'' & ``Lang'' & ``Union'' & ``Optional'' & $p$ \\ 
  \hline
  1 & 1 & 1 & 1 & 1 & 0.012 \\ 
  1 & 1 & 1 & 1 & 2 & $1.4 \cdot 10^{-09}$ \\ 
  1 & 1 & 1 & 2 & 1 & $3.1 \cdot 10^{-09}$ \\ 
  1 & 1 & 1 & 2 & 2 & $6.1 \cdot 10^{-11}$ \\ 
  1 & 1 & 2 & 1 & 1 & $3.3 \cdot 10^{-06}$ \\ 
  1 & 1 & 2 & 1 & 2 & $2.7 \cdot 10^{-09}$ \\ 
  1 & 1 & 2 & 2 & 1 & $2 \cdot 10^{-06}$ \\ 
  1 & 1 & 2 & 2 & 2 & $3.6 \cdot 10^{-10}$ \\ 
  1 & 2 & 1 & 1 & 1 & 0.014 \\ 
  1 & 2 & 1 & 1 & 2 & $1.2 \cdot 10^{-10}$ \\ 
  1 & 2 & 1 & 2 & 1 & $2.8 \cdot 10^{-14}$ \\ 
  1 & 2 & 1 & 2 & 2 & $4.1 \cdot 10^{-15}$ \\ 
  1 & 2 & 2 & 1 & 1 & $2.1 \cdot 10^{-05}$ \\ 
  1 & 2 & 2 & 1 & 2 & $2.7 \cdot 10^{-07}$ \\ 
  1 & 2 & 2 & 2 & 1 & 0.0072 \\ 
  1 & 2 & 2 & 2 & 2 & $1.6 \cdot 10^{-05}$ \\ 
  2 & 1 & 1 & 1 & 1 & 0.28 \\ 
  2 & 1 & 1 & 1 & 2 & $3 \cdot 10^{-07}$ \\ 
  2 & 1 & 1 & 2 & 1 & $3.3 \cdot 10^{-07}$ \\ 
  2 & 1 & 1 & 2 & 2 & $1.7 \cdot 10^{-08}$ \\ 
  2 & 1 & 2 & 1 & 1 & 0.0023 \\ 
  2 & 1 & 2 & 1 & 2 & $6.5 \cdot 10^{-07}$ \\ 
  2 & 1 & 2 & 2 & 1 & 0.00032 \\ 
  2 & 1 & 2 & 2 & 2 & $1.3 \cdot 10^{-06}$ \\ 
  2 & 2 & 1 & 1 & 1 & 0.013 \\ 
  2 & 2 & 1 & 1 & 2 & $1 \cdot 10^{-11}$ \\ 
  2 & 2 & 1 & 2 & 1 & $3.8 \cdot 10^{-11}$ \\ 
  2 & 2 & 1 & 2 & 2 & $4.1 \cdot 10^{-15}$ \\ 
  2 & 2 & 2 & 1 & 1 &   1 \\ 
  2 & 2 & 2 & 1 & 2 & $2.3 \cdot 10^{-05}$ \\ 
  2 & 2 & 2 & 2 & 1 &   1 \\ 
  2 & 2 & 2 & 2 & 2 & 0.99 \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{figure}[ht!]
  \centerline{%
    \includegraphics[width=.9\textwidth]{frac32normal.pdf}}
  \caption{A normal plot of the Fractional Factorial Experiment with
    32 runs. The labelled points are considered significant by using
    the Lenth criterion at a level
    $\alpha=0.05$.}
  \label{fig:frac32normal}
\end{figure}

@@

\begin{figure}[ht!]
  \centerline{%
    \includegraphics[width=0.96\textwidth]{frac64normal.pdf}}
  \caption{A normal plot of the Fractional Factorial Experiment with
    64 runs. The labelled points are considered significant by using
    the Lenth criterion. Triangles correspond to a level $\alpha=0.15$
    and crosses at $\alpha=0.05$. The use of different symbols have
    been added to FrF2 by the authors.}\label{fig:frac64normal}
\end{figure}

@@

\begin{figure}[ht!]
  \centerline{%
  \includegraphics[width=.9\textwidth]{frac64hugenormal.pdf}}
  \caption{A normal plot of the Fractional Factorial Experiment with
    64 runs where the level 2 of ``TripleC'' is 15 MTriples. The
    labelled points are considered significant by using the Lenth
    criterion at a level $\alpha=0.05$.}\label{fig:frac64hugenormal}
\end{figure}

@@

%\bibliographystyle{plain}
%\bibliographystyle{abbrv}
%\bibliographystyle{jbact}
%\bibliographystyle{splncs03}
%\bibliography{practical,selectivity,federation,benchmarks,optimization,experimentsperformance,specs}
\input{iswc2013.bbl}
\end{document}
